Rule Synthesis for Rule-based Intrusion Detectors

### 2020-05-25 paper/paper.tex
fix on related works
pass related work

\noindent
\textbf{Alternative methods for rule synthesis.} We discuss in the
following techniques similar in goal to ours. Unfortunately, we could
not access the tools implementing the techniques for side
comparison. Nevertheless, we found that there are important
differences between our work and related work. Vollmer
\etal{}~\cite{vollmer-etal-cics2011} propose the use of search to
derive rules for Snort, which are very similar to \suri's.  \Fix{Their
approach focuses on icmp attacks.} \Gui{The dataset used by them consists of manually produced ICMP packets using packet creation tools.}. It models every field of the icmp
packet in the chromosome of candidate solutions (\ie{}, a rule) of a
genetic search. The initial population is a set of candidate solutions
that randomly combine options, modelling the packet fields. They show
that they were able to generate rules without false negatives and only
3 false positives across 33K benign packets. The key differences
between our work and Vollmer~\etal{}'s are 1) the support of different
protocols and 2) the use of ranking to discriminate alternative
plausible rules. As for 1, our approach supports a variety of
protocols and options that refer to message payloads which
significantly increase the search space. The icmp rules they generate only
refer to message fields (see Section~\ref{sec:rules-and-packets}).
\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a secao 2.3 
diz q \textbf{geralmente} as regras nao-http possuem apenas field options}. }
As for 2, our solution additionally ranks plausible options, which
contain \emph{no} false positives or negatives with respect to the
benign and mallicious traffic considered.\Gui{Guruprasad \cite{guruprasad} also proposes an evolutionary framework, creating their own ICMP dataset as Vollmer, but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks. The produced rules reported a detection rate up to 96\%. Although Guruprasad supports more protocols than Vollmer, our approach still supports more options and protocols, making both key differences considered when comparing with Vollmer also relevant with Guruprasad.} Muthuregunathan
\etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
create rules. They characterize candidate solutions as a set of field
values and use a clustering algorithm---hybrid k-means partitioned
around menoids---to isolate malicious from benign
traffic. Subsequently, they optimize the formed clusters using genetic
algorithms and hill-climbing search. Ou approach differs from theirs
in the method used, results, and support for content
options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
to synthesize http rules that take malicious traffic as input. ARGS
looks for a common pattern across the different inputs to define
rules. It does not take existing rules as input or benign
traffic. Consequently, false positive can be produced and need to be
handled manually by users. Our approach, in contrast, does not
generate false positives w.r.t. the input traffic. \Gui{Fallahi and colleagues proposes a flow-based solution for automated rule generation. Instead of analyzing the packets individually, their algorithm investigates and creates the rules for flows in the network. The algorithm was tested using the ISCX 2012 dataset and showed good results in terms of precision, but it is limited to the type of attack. As it won't analyse individual details of each packet, it wouldn't be able to identify and generate rules for attacks that require detailed options, such as the HTTP PESC, RCE and LFI. https://ieeexplore.ieee.org/abstract/document/7585840}

@@

\noindent
\textbf{Rule repair.}~Hallahan~\etal{}~\cite{8102263} propose to
repair firewall rules based on false positive detected by the
rules. Although firewall and \nids\ rules have differences and the
problem addressed is different---repair not synthesis---the work is
similar in spirit and inspiring. We remain to evaluate how \tname\ can
be used for repair. \Gui{Ganesan and colleagues introduces an idea of generating new sets of rules based on modifications of existing rules. They used the community edition of snort rules from 2012 for the modifications and tested with the the pcap file from the MACCDC 2012 dataset. It was reported an increase in the number of alerts, but it wasn't analysed the false alarmrates for abduced rules. Their technique is parallel to the rule creation strategies, as the objective of the modified rules is to capture false positives, not new attacks. https://arxiv.org/abs/1903.12101 \Fix{I'm not sure if we can consider it 'rule repair', but I don't know where to put it.}}

>>>>

\noindent
\textbf{Alternative methods for rule synthesis.} We discuss in the
following techniques similar in goal to ours. Unfortunately, we could
not access the tools implementing the techniques for side
comparison. Nevertheless, we found that there are important
differences between our work and related work. Vollmer
\etal{}~\cite{vollmer-etal-cics2011} propose the use of search to
derive rules for Snort, which are very similar to \suri's.  \Fix{Their
approach focuses on icmp attacks.} \Gui{The dataset used by them consists of manually produced ICMP packets using packet creation tools.}. It models every field of the icmp
packet in the chromosome of candidate solutions (\ie{}, a rule) of a
genetic search. The initial population is a set of candidate solutions
that randomly combine options, modelling the packet fields. They show
that they were able to generate rules without false negatives and only
3 false positives across 33K benign packets. The key differences
between our work and Vollmer~\etal{}'s are 1) the support of different
protocols and 2) the use of ranking to discriminate alternative
plausible rules. As for 1, our approach supports a variety of
protocols and options that refer to message payloads which
significantly increase the search space. The icmp rules they generate only
refer to message fields (see Section~\ref{sec:rules-and-packets}).
\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a secao 2.3 
diz q \textbf{geralmente} as regras nao-http possuem apenas field options}. }
As for 2, our solution additionally ranks plausible options, which
contain \emph{no} false positives or negatives with respect to the
benign and mallicious traffic considered.

\Gui{Guruprasad \etal{}\cite{guruprasad} also proposes an evolutionary framework, creating their own ICMP dataset as Vollmer, but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks. The produced rules reported a detection rate up to 96\%. Although Guruprasad supports more protocols than Vollmer, our approach still supports more options and protocols, making both key differences considered when comparing with Vollmer also relevant with Guruprasad.}

Muthuregunathan \etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
create rules. They characterize candidate solutions as a set of field
values and use a clustering algorithm---hybrid k-means partitioned
around menoids---to isolate malicious from benign
traffic. Subsequently, they optimize the formed clusters using genetic
algorithms and hill-climbing search. Ou approach differs from theirs
in the method used, results, and support for content
options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
to synthesize http rules that take malicious traffic as input. ARGS
looks for a common pattern across the different inputs to define
rules. It does not take existing rules as input or benign
traffic. Consequently, false positive can be produced and need to be
handled manually by users. Our approach, in contrast, does not
generate false positives w.r.t. the input traffic.

\Gui{Fallahi \etal{}\cite{fallahi} proposes a flow-based solution for automated rule generation. Instead of analyzing the packets individually, their algorithm investigates and creates the rules for flows in the network. The algorithm was tested using the ISCX 2012 dataset and showed good results in terms of precision, but it is limited to the type of attack. As it won't analyse individual details of each packet, it wouldn't be able to identify and generate rules for attacks that require detailed options, such as the HTTP PESC, RCE and LFI.}

@@

\noindent
\textbf{Rule repair.}~Hallahan~\etal{}~\cite{8102263} propose to
repair firewall rules based on false positive detected by the
rules. Although firewall and \nids\ rules have differences and the
problem addressed is different---repair not synthesis---the work is
similar in spirit and inspiring. We remain to evaluate how \tname\ can
be used for repair.

\Gui{Ganesan \etal{} introduces an idea of generating new sets of rules based on modifications of existing rules. They used the community edition of snort rules from 2012 for the modifications and tested with the the pcap file from the MACCDC 2012 dataset. It was reported an increase in the number of alerts, but it wasn't analysed the false alarmrates for abduced rules. Their technique is parallel to the rule creation strategies, as the objective of the modified rules is to capture variations and evolutions of attacks.}
